---
title: "Assignment-03"
author: "Emma Bonanno", "Mike Lidwin", "Kayla Myros"
date: "9/23/2021"
output: html_document
---

```{r load packages, message=FALSE}
#above line labels the code chunk as "load packages" without seeing any of the package loading messages
  library(sf)
  library(tidyverse)
  library(ggthemes)
  library(ggspatial)
  library(units)
  library(nngeo)
```

```{r load data, message=FALSE}
bshop <- st_read("point-a-bshop.kml", 
                  quiet = TRUE) 

library <- st_read("point-b-libraries.kml", 
                 quiet = TRUE)

food <- st_read("polygon-c-food-access.kml", 
                   quiet = TRUE) 

clusters <- st_read("polygon-d-clusters.kml", 
                 quiet = TRUE) 
```
## Transforming the Data to the Washington DC State Plane

```{r}
DC_state_plane <- "+proj=lcc +lat_1=38.3 +lat_2=39.45 +lat_0=37.66666666666666 +lon_0=-77 +x_0=400000 +y_0=0 +ellps=GRS80 +units=m +no_def"

bshop <- bshop %>%
  st_transform(DC_state_plane)

food <- food %>%
  st_transform(DC_state_plane)

clusters <- clusters %>%
  st_transform(DC_state_plane)

library <- library %>%
  st_transform(DC_state_plane)
```

```{r section, message=FALSE}
###KAYLA BELOW
```



```{r section, message=FALSE}
###EMMA BELOW
```
## How many Books and Barbershops are in each neighborhood in Washington, DC?

This map shows the number of locations of Books and Barbershops programs in each neighborhood in DC. The goal was to assess if this particular program was particularly over- or under- represented in certain neighborhoods. 
```{r}
clusters <- clusters %>%
  mutate(num_bshop = lengths(st_covers(clusters, bshop)))

ggplot(clusters) +
  geom_sf(color = "white", 
          aes(fill = num_bshop)) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Books and Barbershops Locations
in DC Neighborhoods",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.position = "left",
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```
```{r}
```
## How many Books and Barbershops are in each Low Food Access Area in DC?

The Books and Barbershops program is an attempt to raise awareness of the value of academic mentoring and early childhood literacy, both basic community needs. We chose to look at the locations of these data points in reference to a data set that demonstrates unequal access to another basic need - food.   
```{r}
food <- food %>%
  mutate(num_bshop = lengths(st_covers(food,bshop)))

ggplot(food) +
  geom_sf(color = NA, 
          aes(fill = num_bshop)) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Books and Barbershops in
Low Food Access Areas
",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.position = "left", legend.background = element_rect(fill = alpha("white", 0.5),  color = "gray"))
```

```{r section, message=FALSE}
###MIKE BELOW
```


```{r}
```
## Contribution Statement

We met as a team to pick our datasets and decide which combinations of points and polygons to map. Individually, we each created 2 of the 6 maps, and we helped each other troubleshoot throughout the process. Mike created a shared GitHub repo, Emma wrote the contribution statement, and Kayla did a final check through of the RMD file for cohesion.

Emma - 10 points
Kayla - 10 points
Mike - 10 points

