---
title: "Assignment-03"
author: "Emma Bonanno", "Mike Lidwin", "Kayla Myros"
date: "9/23/2021"
output: html_document
---

```{r load packages, message=FALSE}
#above line labels the code chunk as "load packages" without seeing any of the package loading messages
  library(sf)
  library(tidyverse)
  library(ggthemes)
  library(ggspatial)
  library(units)
  library(nngeo)
```

```{r load data, message=FALSE}
bshop <- st_read("point-a-bshop.kml", 
                  quiet = TRUE) 

library <- st_read("point-b-libraries.kml", 
                 quiet = TRUE)

food <- st_read("polygon-c-food-access.kml", 
                   quiet = TRUE) 

clusters <- st_read("polygon-d-clusters.kml", 
                 quiet = TRUE) 
```
## Transforming the Data to the Washington DC State Plane

```{r}
DC_state_plane <- "+proj=lcc +lat_1=38.3 +lat_2=39.45 +lat_0=37.66666666666666 +lon_0=-77 +x_0=400000 +y_0=0 +ellps=GRS80 +units=m +no_def"

bshop <- bshop %>%
  st_transform(DC_state_plane)

food <- food %>%
  st_transform(DC_state_plane)

clusters <- clusters %>%
  st_transform(DC_state_plane)

library <- library %>%
  st_transform(DC_state_plane)
```

```{r section, message=FALSE}
###KAYLA BELOW
```

##Buffer the library data points with a half kilometer buffer
##Create a variable to find the overlap of Barbershop Books within a half kilometer of a public library
##Merge the library file with the Barbershop Books file 

```{r}
libr_buffer <- st_buffer(library, dist = 500) %>%
  st_union()

bnb_libr <- bshop[libr_buffer,]

bshop <- bshop %>%
  st_join(bnb_libr) %>%
  mutate(by_library = !is.na(Name.y))


n_libr_bnb <- sum(bshop$by_library)

n_bnb <- length(bshop$by_library)

pct_library_bshop <- n_libr_bnb / n_bnb



```

##Mapping the number and proportion of Book-Barbershop points within a half kilometer of a public library point.

Barbershop Books is a community literacy program that provides age appropriate reading materials for elementary school aged children in barbershops around D.C. As of 2021, there were 30 recorded barbershops participating in this program. The map below compares the number of these participating barbershops within a half kilometer distance to a public library. This comparison provides context for areas in D.C. that have access to free reading materials. 

```{r}
left_side  <- st_bbox(clusters)$xmin
top_side <- st_bbox(clusters)$ymax

# ggplot(libr_buffer) +
#   geom_sf(data = clusters, lty=0.2)+
#   geom_sf(alpha = 0.25)+
#   # geom_sf(data = bshop, color= "navy", alpha =.75)+
#   geom_sf(data = bshop, size = 1, alpha =.5,
#            aes(color = "red"))+


ggplot(clusters) +
  geom_sf(fill = NA, color= "darkgray" )+
  geom_sf(data = libr_buffer, color= "lightgray", alpha =.75,
          aes(fill = ""))+
  geom_sf(data = bshop, size = 1, alpha =.5,
           aes(color = "navy"))+
  scale_color_manual(values = c( "navy"),
          name = "Barbershops Books \nby proximity to Library", 
          labels = c("barbershops")) +
  scale_fill_manual(values = c("lightgray"),name ="Library Buffer")+
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_bnb, big.mark = ","),
                         " Barbershop Books locations in D.C. \n", 
                         prettyNum(n_libr_bnb, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_library_bshop, digits = 0),
                         "%) are within 0.5\n kilometer of a public.",
                         sep = ""),
           hjust = 0, vjust = .5, size = 2) +
  theme_map() +
  theme(panel.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                        color = "gray"))
```


##The average density of public library points in each neighborhood cluster polygon

For the second analysis, we mapped the density of libraries in the D.C. neighborhood clusters. This analysis shows which neighborhood clusters are served or under-served by libraries. 



```{r}
clusters <- clusters %>%
  mutate(num_library = lengths(st_covers(clusters, library)))


clusters <- clusters %>%
  mutate(area = set_units(st_area(clusters), km^2)) %>%
  mutate(library_dens = as.numeric(num_library / area))

##create map##

ggplot(clusters) +
  geom_sf(color = "white", 
          aes(fill = num_library)) +
  scale_fill_viridis_c(option = "plasma", name = "D.C neighborhoods\nby library density",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "library / sq km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```





```{r section, message=FALSE}
###EMMA BELOW
```
## How many Books and Barbershops are in each neighborhood in Washington, DC?

This map shows the number of locations of Books and Barbershops programs in each neighborhood in DC. The goal was to assess if this particular program was particularly over- or under- represented in certain neighborhoods. 
```{r}
clusters <- clusters %>%
  mutate(num_bshop = lengths(st_covers(clusters, bshop)))

ggplot(clusters) +
  geom_sf(color = "white", 
          aes(fill = num_bshop)) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Books and Barbershops Locations
in DC Neighborhoods",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.position = "left",
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```
```{r}
```
## How many Books and Barbershops are in each Low Food Access Area in DC?

The Books and Barbershops program is an attempt to raise awareness of the value of academic mentoring and early childhood literacy, both basic community needs. We chose to look at the locations of these data points in reference to a data set that demonstrates unequal access to another basic need - food.   
```{r}
food <- food %>%
  mutate(num_bshop = lengths(st_covers(food,bshop)))

ggplot(food) +
  geom_sf(color = NA, 
          aes(fill = num_bshop)) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Books and Barbershops in
Low Food Access Areas
",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.position = "left", legend.background = element_rect(fill = alpha("white", 0.5),  color = "gray"))
```

```{r section, message=FALSE}
###MIKE BELOW
```


```{r}
```
## Contribution Statement

We met as a team to pick our datasets and decide which combinations of points and polygons to map. Individually, we each created 2 of the 6 maps, and we helped each other troubleshoot throughout the process. Mike created a shared GitHub repo, Emma wrote the contribution statement, and Kayla did a final check through of the RMD file for cohesion.

Emma - 10 points
Kayla - 10 points
Mike - 10 points

