---
title: "Washington DC maps"
author: "Emma Bonanno"
date: "9/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)

```
Load Data

```{r}

bshop <- st_read("C:/Users/emmab/Desktop/Books.kml", quiet = TRUE)
food <- st_read("C:/Users/emmab/Desktop/Food.kml", quiet = TRUE)
clusters <- st_read("C:/Users/emmab/Desktop/Nhoods.kml", quiet = TRUE)
library <- st_read("C:/Users/emmab/Desktop/Libraries.kml", quiet = TRUE) 

```

## Transforming the Data

```{r}
DC_state_plane <- "+proj=lcc +lat_1=38.3 +lat_2=39.45 +lat_0=37.66666666666666 +lon_0=-77 +x_0=400000 +y_0=0 +ellps=GRS80 +units=m +no_def"

bshop <- bshop %>%
  st_transform(DC_state_plane)

food <- food %>%
  st_transform(DC_state_plane)

clusters <- clusters %>%
  st_transform(DC_state_plane)

library <- library %>%
  st_transform(DC_state_plane)
```
Maps
```{r}
ggplot(food) +
  geom_sf(fill = "orange", color = NA) +
  geom_sf(data = bshop, size = 1) +
  geom_sf(data= clusters, fill = NA, color = "gray") +
  theme_map() +
  annotation_scale()
```
```{r}
```
1000 km radius around Libraries

```{r}
library_buffer <- st_buffer(library, dist = 1000) %>%
  st_union()

ggplot(library_buffer) +
  geom_sf() +
  theme_map()
```
Books and Barbershops within a 1000 m radius of Libraries
```{r}
bshop_library <- bshop[library_buffer,]
  
ggplot(library_buffer) +
  geom_sf() +
  geom_sf(data = bshop_library, 
          color = "darkgreen", 
          size = 2) +
  theme_map()
```
Join dataframes - How many bshops are within 1000 m of a library?
```{r}
bshop <- bshop %>%
  st_join(bshop_library) %>%
  mutate(by_library = !is.na(Name.y))

n_library_bshop <- sum(bshop$by_library)

n_library_bshop
```
What percent of all bshops is this?
```{r}
n_bshop <- length(bshop$by_library)

pct_library_bshop <- n_library_bshop / n_bshop

pct_library_bshop

```







How many Bshops are in each neighborhood?
```{r}
clusters <- clusters %>%
  mutate(num_bshop = lengths(st_covers(clusters, bshop)))

ggplot(clusters) +
  geom_sf(color = "white", 
          aes(fill = num_bshop)) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Books and Barbershops Locations
in DC Neighborhoods",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.position = "left",
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```
Number of bshops in food
```{r}
food <- food %>%
  mutate(num_bshop = lengths(st_covers(food,bshop)))

ggplot(food) +
  geom_sf(color = NA, 
          aes(fill = num_bshop)) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Books and Barbershops in 
Low Food Access Areas
  ",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.position = "left", legend.background = element_rect(fill = alpha("white", 0.5),  color = "gray"))
```

Density of Books and Barbershops in Low Access Food Areas
```{r}
food <- food %>%
  mutate(area = set_units(st_area(food), km^2)) %>%
  mutate(bshop_dens = as.numeric(num_bshop / area))

ggplot(food) +
  geom_sf(color = NA,
          aes(fill = bshop_dens)) +
    scale_fill_viridis_c(name = 
                           "Low Food Access Areas by Number of Books and Barbershops",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Books and Barbershops per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```
Number of Libraries in Low Food Access
```{r}
food <- food %>%
  mutate(num_library = lengths(st_covers(food,library)))

ggplot(food) +
  geom_sf(color = NA, 
          aes(fill = num_library)) +
  scale_fill_viridis_c(name = "Number of Libraries in Low Food Access Areas",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Libraries")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5),  color = "gray"))
```



Density of Libraries in Low Food Access
```{r}

food <- food %>%
  mutate(num_library = lengths(st_covers(food, library))) %>%
  mutate(area = set_units(st_area(food), km^2)) %>%
  mutate(libary_dens = as.numeric(num_library / area))

ggplot(food) +
  geom_sf(color = NA,
          aes(fill = library_dens)) +
    scale_fill_viridis_c(name = 
                           "Density of Libraries in Low Food Access Areas",
                       breaks = breaks <- seq(0, 3, by = .5),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Libraries per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```



















